Mental Health Plays Significant Role in Physical Wellness From 1972 to 2022 US General Social Surveys

### 2024-04-23 paper/paper.qmd
update graph formatting and descriptions

```{r}
#| label: tbl-sub-vars
#| tbl-cap: Overview of the Mental and Physical Unwellness
#| echo: false
#| warning: false
#| message: false

@@

kable(head(days_data), 
      caption = "Counted days of Mental unwellness with General Health Status", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

@@

The survey queried participants with questions, such as "Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?" and "Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good?" and expecting a response of the number of days in a month. The details of the questions used are listed in [Appendix -@sec-ques-details]. Significantly, between 1998 and 2022, the structure of the questions has consistently remained the same. This consistency in formatting ensures the data's stability, facilitating a high level of consistency for thorough analysis. The data consists of a combination of in-person, telephone, and web surveys.

@@

```{r}
#| label: tbl-demo-overview
#| tbl-cap: Overview of the Demographic Data 
#| echo: false
#| warning: false
#| message: false

@@

kable(head(demo_data), 
      caption = "Counted days of Physical and Mental unwellness of Population by Age and Gender Groups", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

@@

One of the key variables used for further analysis is whether the respondent has been diagnosed as having depression in the past. The survey used the question "Now I would like to ask you some questions about general health conditions. Has a doctor, nurse, or other health professional EVER told you that you had: D. Depression?" to collect a categorical response of whether "Yes" or "No". The variable is an external indication of respondents' mental health status, accompanying the main counted days variable that is used for the analysis (@tbl-analysis-vars). About 21% of the participants have been diagnosed as having depression in the past. According to the Morbidity and Mortality Weekly Report from Centers for Diseases Control and Prevention (CDC) in U.S.[@CDControlPrevention_2023], 18.4% of U.S. adults reported having ever been diagnosed with depression during 2020. The percentage in dataset is slightly higher than weekly reports, which could be a result from error, sampling differences, and other factors.

```{r}
#| label: tbl-analysis-vars
#| tbl-cap: Overview of the Main Analysis Dataset
#| echo: false
#| warning: false
#| message: false

@@

kable(head(analysis_data), 
      caption = "Counted days of Physical and Mental unwellness with Depression Diagnosis", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)

per <- round(count(filter(analysis_data,depress=="Yes")) / count(analysis_data), 2) * 100
```

@@

A variable collecting general health status is furthermore included for additional analyses (@tbl-health-vars). The variable is used as a supplementary tool indicating respondents' health in general to assist the analysis result. The question consists of four categorical responses asking the respondents: "Would you say your own health, in general, is excellent, good, fair, or poor?" The respondents are expected to choose one of the options to answer.

@@

vars <- c("ID","Mental Unwellness","General Health")

kable(head(mentalVsHealth), 
      caption = "Counted days of Mental unwellness with General Health Status", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

@@

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

### Diagnostics

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

>>>>

```{r}
#| label: tbl-sub-vars
#| tbl-cap: Counted days of Mental unwellness with General Health Status
#| echo: false
#| warning: false
#| message: false

@@

kable(head(days_data), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

@@

The survey queried participants with questions, such as "Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?" and "Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good?" and expecting a response of the number of days in a month. Further details of the questions used are listed in [Appendix -@sec-ques-details]. Significantly, between 1998 and 2022, the structure of the questions has consistently remained the same. This consistency in formatting ensures the data's stability, facilitating a high level of consistency for thorough analysis. The data consists of a combination of in-person, telephone, and web surveys.

@@

```{r}
#| label: tbl-demo-overview
#| tbl-cap: Counted days of Physical and Mental unwellness of Population by Age and Gender Groups
#| echo: false
#| warning: false
#| message: false

@@

kable(head(demo_data), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

@@

One of the key variables used for further analysis is whether the respondent has been diagnosed as having depression in the past. The survey used the question "Now I would like to ask you some questions about general health conditions. Has a doctor, nurse, or other health professional EVER told you that you had: D. Depression?" to collect a categorical response of whether "Yes" or "No". The variable is an external indication of respondents' mental health status, accompanying the main counted days variable that is used for the analysis (@tbl-analysis-vars). From @tbl-analysis-depress, about 21% of the participants have been diagnosed as having depression in the past. According to the Morbidity and Mortality Weekly Report from Centers for Diseases Control and Prevention (CDC) in U.S.[@CDControlPrevention_2023], 18.4% of U.S. adults reported having ever been diagnosed with depression during 2020. The percentage in dataset is slightly higher than weekly reports, which could be a result from error, sampling differences, and other factors.

```{r}
#| label: tbl-analysis-vars
#| tbl-cap: Counted days of Physical and Mental unwellness with Depression Diagnosis
#| echo: false
#| warning: false
#| message: false

@@

kable(head(analysis_data), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)

per <- round(count(filter(analysis_data,depress=="Yes")) / count(analysis_data), 2) * 100

```
\break
```{r}
#| label: tbl-analysis-depress
#| tbl-cap: Distribution of Depression Diagnosis Responses
#| echo: false
#| warning: false
#| message: false

depresstable <- count(analysis_data, depress)
total <- c("Total",nrow(analysis_data))
depresstable <- rbind(depresstable ,total)
kable(depresstable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Responses", "Occurrences"))
```

@@

A variable collecting general health status is furthermore included for additional analyses (@tbl-health-vars). The variable is used as a supplementary tool indicating respondents' health in general to assist the analysis result. The question consists of four categorical responses asking the respondents: "Would you say your own health, in general, is excellent, good, fair, or poor?" The respondents are expected to choose one of the options to answer. In @tbl-health-rep, number of occurrences for each responses are counted. Majority of the repspondents reported to have good (51%) or excellent (29%) health status.

@@

per_good <- round(nrow(filter(mentalVsHealth, health == 2))/nrow(mentalVsHealth),2) * 100
per_exc <- round(nrow(filter(mentalVsHealth, health == 1))/nrow(mentalVsHealth),2) * 100

vars <- c("ID","Mental Unwellness","General Health")
kable(head(mentalVsHealth), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)

```{r}
#| label: tbl-health-rep
#| tbl-cap: Distribution of General Health Responses
#| echo: false
#| warning: false
#| message: false
healthtable <- count(mentalVsHealth, health)
total <- c("Total",nrow(mentalVsHealth))
healthtable <- rbind(healthtable ,total)
status <- data.frame(c("Excellent", "Good", "Fair", "Poor","Total"))
colnames(status) <- c("Health Status")
healthtable <- cbind(status, healthtable)
kable(healthtable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Health Status", "Responses", "Occurrences"))
```

@@

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
\break
### Diagnostics

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "The convergence of the MCMC algorithm is checked for the model."
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
